### 2018-07-05 submission/manuscript.Rmd
change title
update methods

# Assessing Differences in High Fidelity DNA Polymerase for 16S rRNA Gene Sequencing

@@

***Analysis workflow:***
To adjust for unequal sequencing, all samples were rarefied to 1000 sequences for downstream analysis. The total number of OTUs was analyzed for both the fecal and mock community samples. For fecal samples, cycle dependent affects on Bray-Curtis indices were assessed for cycle group and within individual differences from the previous cycle (e.g. 20 versus 25, 25 versus 30). Based on these observations we analyzed potential reasons for these differences using the mock community samples. Analysis of the mock community of each polymerase for sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras before and after chimera removal with VSEARCH was assessed. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed.

***Statistical analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data transformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. Differences in the total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the comparison in fecal samples the data was normalized to each individual by cycle number to account for the biological variation between people. Bray-Curtis distance matrices were generated using mothur. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and Kruskal-Wallis tests within R. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable, correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

>>>>

# Assessing the Differences in 16S rRNA Gene Sequencing Due to High Fidelity DNA Polymerase

@@

***Analysis workflow:***
To adjust for unequal sequencing, all samples were rarefied to 1000 sequences for downstream analysis. Analysis of the mock community of each polymerase for GC-based amplification differences, sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras before and after chimera removal with VSEARCH was assessed. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed. The total number of OTUs, taxa differences, and Bray-Curtis indices were analyzed for both the fecal and mock community samples. Finally, Random Forest models were created to assess whether classification of polymerase, cycles, or individual could be performed best using 16S rRNA gene sequencing data. Additionally, overlap between the most important OTUs to the three models was assessed using mean decrease in accuracy (MDA).

***Statistical analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data transformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. Differences in the total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the comparison of the number of OTUS in fecal samples the data was normalized to each individual by cycle number to account for the biological variation. Bray-Curtis distance matrices were generated using mothur. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and Kruskal-Wallis tests within R. The Random Forest models were run using the caret package (v `r paste(packageVersion("caret")[1])`) [@caret_citation]. A total of 100 10-fold CV runs on different 80/20 splits of the data was run to generate a range of the Logloss value. The probability of a correct call was obtained from this Logloss value by taking the negative natural logrithim. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable, correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

### 2018-07-05 submission/manuscript.tex
update methods

\textbf{\emph{Analysis workflow:}} To adjust for unequal sequencing, all
samples were rarefied to 1000 sequences for downstream analysis. The
total number of OTUs was analyzed for both the fecal and mock community
samples. For fecal samples, cycle dependent affects on Bray-Curtis
indices were assessed for cycle group and within individual differences
from the previous cycle (e.g.~20 versus 25, 25 versus 30). Based on
these observations we analyzed potential reasons for these differences
using the mock community samples. Analysis of the mock community of each
polymerase for sequence error rate, number of sequences with an error,
base substitution, and numbers of chimeras before and after chimera
removal with VSEARCH was assessed. Additionally, the correlation between
the number of chimeras and the number of OTUs was also assessed.

\textbf{\emph{Statistical analysis:}} All analysis was done with the R
(v 3.4.4) software package (R Core Team, 2017). Data transformation and
graphing was completed using the tidyverse package (v 1.2.1) and colors
selected using the viridis package (v 0.4.1) (Garnier, 2017; Wickham,
2017). Differences in the total number of OTUs were analyzed using an
ANOVA with a tukey post-hoc test. For the comparison in fecal samples
the data was normalized to each individual by cycle number to account
for the biological variation between people. Bray-Curtis distance
matrices were generated using mothur. The distance matrix data was
analyzed using PERMANOVA with the vegan package (v 2.4.5) (Oksanen et
al., 2017) and Kruskal-Wallis tests within R. For both error and chimera
analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc
test. Where applicable, correction for multiple comparison utilized the
Benjamini-Hochberg method (Benjamini \& Hochberg, 1995).

@@

>>>>

\textbf{\emph{Analysis workflow:}} To adjust for unequal sequencing, all
samples were rarefied to 1000 sequences for downstream analysis.
Analysis of the mock community of each polymerase for GC-based
amplification differences, sequence error rate, number of sequences with
an error, base substitution, and numbers of chimeras before and after
chimera removal with VSEARCH was assessed. Additionally, the correlation
between the number of chimeras and the number of OTUs was also assessed.
The total number of OTUs, taxa differences, and Bray-Curtis indices were
analyzed for both the fecal and mock community samples. Finally, Random
Forest models were created to assess whether classification of
polymerase, cycles, or individual could be performed best using 16S rRNA
gene sequencing data. Additionally, overlap between the most important
OTUs to the three models was assessed using mean decrease in accuracy
(MDA).

\textbf{\emph{Statistical analysis:}} All analysis was done with the R
(v 3.4.4) software package (R Core Team, 2017). Data transformation and
graphing was completed using the tidyverse package (v 1.2.1) and colors
selected using the viridis package (v 0.4.1) (Garnier, 2017; Wickham,
2017). Differences in the total number of OTUs were analyzed using an
ANOVA with a tukey post-hoc test. For the comparison of the number of
OTUS in fecal samples the data was normalized to each individual by
cycle number to account for the biological variation. Bray-Curtis
distance matrices were generated using mothur. The distance matrix data
was analyzed using PERMANOVA with the vegan package (v 2.4.5) (Oksanen
et al., 2017) and Kruskal-Wallis tests within R. The Random Forest
models were run using the caret package (v 6.0.78) (Jed Wing et al.,
2017). A total of 100 10-fold CV runs on different 80/20 splits of the
data was run to generate a range of the Logloss value. The probability
of a correct call was obtained from this Logloss value by taking the
negative natural logrithim. For both error and chimera analysis, samples
were tested using Kruskal-Wallis with a Dunns post-hoc test. Where
applicable, correction for multiple comparison utilized the
Benjamini-Hochberg method (Benjamini \& Hochberg, 1995).

@@

\hypertarget{ref-caret_citation}{}
Jed Wing MKC from., Weston S., Williams A., Keefer C., Engelhardt A.,
Cooper T., Mayer Z., Kenkel B., R Core Team., Benesty M., Lescarbeau R.,
Ziem A., Scrucca L., Tang Y., Candan C., Hunt. T. 2017. \emph{Caret:
Classification and regression training}.
